---
title: "dengue"
author: "Sin Myeong Eun"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
packages = c('rgdal', 'maptools', 'raster','spatstat', 'tmap','readr','dplyr','sf','ggplot2','lubridate')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
```

```{r}
dengue <- read_csv("data/dengue outbreak_Singapore_2020.csv")
dengue_sf <- st_as_sf(dengue, coords = c("longitude", "latitude"), crs=4326) %>%
  st_transform(crs=3414)
# st_crs(dengue_sf)
# st_bbox(dengue_sf)
dengue_sp <- as(dengue_sf, "Spatial")
dengue_sp <- as(dengue_sp, "SpatialPoints")
dengue_ppp <- as(dengue_sp, "ppp")

mpsz <- readOGR(dsn = "data", layer="MP14_SUBZONE_WEB_PL")
# plot(mpsz)

sg <- readOGR(dsn = "data", layer="SG_main")
sg_sp <- as(sg, "SpatialPolygons")
sg_sp <- spTransform(sg_sp, "+init=epsg:3414")
sg_owin <- as(sg_sp, "owin")
# bbox(sg_sp)
# spTransform(sg,3414)
# plot(sg_owin)

dengueSG_ppp = dengue_ppp[sg_owin]
plot(dengueSG_ppp)
```

Quadrat Analysis
1. Chi-squared test of CSR using quadrat counts
The test hypotheses are:
Ho = The distribution of dengue cases are randomly distributed.
H1= The distribution of dengue cases are not randomly distributed.
The 95% confident interval will be used.

```{r}
qt <- quadrat.test(dengueSG_ppp, 
                   nx = 20, ny = 15)
qt
plot(dengueSG_ppp)
plot(qt, add = TRUE, cex =.1)
```
Conclusion from Chi-squared test:
Since p-value is smaller than the significance level of 0.05 (since confidence interval is 0.95), we can reject the null hypothesis. Thus, the distribution of dengue cases are not randomly distributed

```{r}
quadrat.test(dengueSG_ppp, 
             nx = 20, ny = 15,
             method="M",
             nsim=999)
```
Conclusion from Monte Carlo test:
Since p-value is smaller than the significance level of 0.05 (since confidence interval is 0.95), we can reject the null hypothesis. Thus, the distribution of dengue cases are not randomly distributed

Nearest Neighbour Analysis

```{r}
clarkevans.test(dengueSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("two.sided"),
                nsim=99)
```

Conclusion from Clark-Evans test:
Since p-value is smaller than the significance level of 0.05 (since confidence interval is 0.95), we can reject the null hypothesis. Thus, the distribution of dengue cases are not randomly distributed

Kernel Density Estimation
```{r}
# Fixed Bandwidth
kde_dengueSG_bw <- density(dengueSG_ppp, sigma=bw.diggle, edge=TRUE, kernel="gaussian") 
dengueSG_ppp.km <- rescale(dengueSG_ppp, 1000, "km")
kde_dengueSG_bw <- density(dengueSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")

# Adaptive Bandwidth
kde_dengueSG_adaptive <- adaptive.density(dengueSG_ppp.km, method='kernel')

# Fixed vs Adaptive Bandwidth
par(mfrow=c(1,2))
plot(kde_dengueSG_bw, main = "Fixed Bandwidth")
plot(kde_dengueSG_adaptive, main = "Adaptive Bandwidth")
```

```{r}
# Convert KDE output into grid object
gridded_kde_dengueSG_bw <- as.SpatialGridDataFrame.im(kde_dengueSG_bw)
# spplot(gridded_kde_dengueSG_bw,at = seq(0, 1000, by = 100))
spplot(gridded_kde_dengueSG_bw)
kde_dengueSG_bw_raster <- raster(gridded_kde_dengueSG_bw)
projection(kde_dengueSG_bw_raster) <- CRS("+init=EPSG:3414")
# kde_dengueSG_bw_raster

tm_shape(kde_dengueSG_bw_raster) + 
  tm_raster("v",breaks = c(0,50, 100, 150, 200,250,1000,1500,2000,2500)) +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```
Comparing Spatial Point Patterns using KDE
```{r}
mp = mpsz[mpsz@data$PLN_AREA_N == "MARINE PARADE",]
pg = mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
ck = mpsz[mpsz@data$PLN_AREA_N == "CHOA CHU KANG",]
jw = mpsz[mpsz@data$PLN_AREA_N == "JURONG WEST",]

par(mfrow=c(2,2))
plot(mp, main = "Marine Parade")
plot(pg, main = "Punggol")
plot(ck, main = "Choa Chu Kang")
plot(jw, main = "Jurong West")
# Convert spdf to sp
mp_sp = as(mp, "SpatialPolygons")
pg_sp = as(pg, "SpatialPolygons")
ck_sp = as(ck, "SpatialPolygons")
jw_sp = as(jw, "SpatialPolygons")
# create owin object
mp_owin = as(mp_sp, "owin")
pg_owin = as(pg_sp, "owin")
ck_owin = as(ck_sp, "owin")
jw_owin = as(jw_sp, "owin")
# combine dengue points with subzone areas
dengue_mp_ppp = dengue_ppp[mp_owin]
dengue_pg_ppp = dengue_ppp[pg_owin]
dengue_ck_ppp = dengue_ppp[ck_owin]
dengue_jw_ppp = dengue_ppp[jw_owin]
# rescale to km
dengue_mp_ppp.km = rescale(dengue_mp_ppp, 1000, "km")
dengue_pg_ppp.km = rescale(dengue_pg_ppp, 1000, "km")
dengue_ck_ppp.km = rescale(dengue_ck_ppp, 1000, "km")
dengue_jw_ppp.km = rescale(dengue_jw_ppp, 1000, "km")

par(mfrow=c(2,2))
plot(dengue_mp_ppp.km, main="Marine Parade")
plot(dengue_pg_ppp.km, main="Punggol")
plot(dengue_ck_ppp.km, main="Choa Chu Kang")
plot(dengue_jw_ppp.km, main="Jurong West")

```

```{r}
kde_dengue_mp_bw <- density(dengue_mp_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
kde_dengue_pg_bw <- density(dengue_pg_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_dengue_pg_bw)
kde_dengue_ck_bw <- density(dengue_ck_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
kde_dengue_jw_bw <- density(dengue_jw_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian") 
par(mfrow=c(2,2))
plot(kde_dengue_mp_bw, main="Marine Parade")
plot(kde_dengue_pg_bw, main="Punggol")
plot(kde_dengue_ck_bw, main="Choa Chu Kang")
plot(kde_dengue_jw_bw, main="Jurong West")
```
Computing fixed bandwidth KDE
```{r}
kde_dengue_mp_250 <- density(dengue_mp_ppp.km, sigma=0.25, edge=TRUE, kernel="gaussian")
kde_dengue_pg_250 <- density(dengue_pg_ppp.km, sigma=0.25, edge=TRUE, kernel="gaussian")
kde_dengue_ck_250 <- density(dengue_ck_ppp.km, sigma=0.25, edge=TRUE, kernel="gaussian")
kde_dengue_jw_250 <- density(dengue_jw_ppp.km, sigma=0.25, edge=TRUE, kernel="gaussian")
par(mfrow=c(2,2))
plot(kde_dengue_mp_250, main="Marine Parade")
plot(kde_dengue_pg_250, main="Punggol")
plot(kde_dengue_ck_250, main="Choa Chu Kang")
plot(kde_dengue_jw_250, main="Jurong West")
```
Analysing Spatial Point Process Using Nearest Neighbour Index

H0 = The distribution of dengue spread is randomly distributed
H1 = The distribution of dengue spread is not randomly distributed
Confidence interval = 95%
```{r}
# Clark and Evans test: Marine Parade
clarkevans.test(dengue_mp_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```
```{r}
# Clark and Evans test: Punggol
clarkevans.test(dengue_pg_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```
```{r}
# Clark and Evans test: Choa Chu Kang
clarkevans.test(dengue_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```
```{r}
# Clark and Evans test: Jurong West
clarkevans.test(dengue_jw_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```
Since p-value is significantly small at 0.002 and it’s smaller than the significance level of 0.05 (since confidence interval is 0.95), we can reject the null hypothesis. Thus, the distribution of dengue spread is not randomly distributed